Algebraic File Synchronization: Adequacy and Completeness

### 2016-10-15 p2/rec-intro.tex
Rewrote proof for new notation

\noindent
Next, we would now like to prove that $\reca$ and $\recb$ can be applied to the replicas,
that is, without loss of generality,
\[ \reca\aFS_B\neq\fsbroken. \]
So that we can formalize statements needed to prove this,
let us introduce relations that describe under what conditions
sequences of commands work, that is, do not break a filesystem.

>>>>

\noindent
Next, we would now like to prove that $\reca$ and $\recb$ can be applied to the replicas,
that is, without loss of generality,
\[ \reca\aFS_B\neq\fsbroken. \]
So that we can formalize statements needed to prove this,
let us introduce partial orders that describe under what conditions
sequences of commands work, that is, do not break a filesystem.

### 2016-10-15 p2/works-inverse.tex
Rewrote proof for new notation

\begin{mydef}[Sets of sequences]
As we will frequently refer to sets of fequences,
we will use calligraphic letters (e.g. $\sqs{A}$, $\sqs{B}$, $\sqs{C}$ and $\seqset{S}$)
to denote such sets for brevity.
We will use $A\cc\seqset{B}$ to denote the set of sequences $\{A\cc B|B\in\seqset{B}\}$,
and $\Dom{\sqs{A}}$ to denote $\bigcap_{A\in\sqs{A}} \Dom{A}$.
\end{mydef}

@@

It is easy to see that the following corollaries are true:

@@

\begin{mycor}\label{workschained}
% An inference rule in the algebra
$\worksmeqsign$ can be chained:
if $\seqset{B'}\subseteq\seqset{B}$, then
$\worksm{\sqs{A}}{\sqs{B}} \wedge \worksm{\sqs{B'}}{\sqs{C}} \Rightarrow \worksm{\sqs{A}}{\sqs{B}\cup\sqs{C}}$.
\end{mycor}

@@

\noindent
We know that whether commands break a filesystem depends solely on their input and output
types and the node, but not their actual output value.
To represent this, we introduce an equivalence relation between filesystems, commands
and sequences of commands.
%% TODO extending it!
\begin{mydef}[Equivalence by type: $\typeeq$]
$ $ % otherwise first list element starts on same line
\begin{itemize}
\item For two filesystems $\FS\typeeq\GS$ iff they are both broken, or they share the same nodes $\setn$,
and for all $n\in\setn$, $\FS(n) \typeeq \GS(n)$.
\item For two commands $\cxynv\typeeq\czwnv$ iff 
$x\typeeq z$ and $y\typeeq w$ and $n=m$.
\item For two sequences of commands $\emptyseq\typeeq\emptyseq$, and
if $A'=\alpha\cc A$ and $B'=\beta\cc B$, then $A'\typeeq B'$ iff $\alpha\typeeq\beta$ and $A\typeeq B$.
\end{itemize}
\end{mydef}

@@

\begin{mydef}[$\T$]
We use $\T$ and write $X\T$ to denote the canonical mapping of 
values, filesystems, commands and sequences to the
representatives of their $\typeeq$-equivalence class.
\begin{itemize}
\item For values, $\vald\T=\vald$, $\valf\T=\fil$ and $\empt\T=\empt$.
\item For filesystems, $\fsbroken\T=\fsbroken$ and in other cases $\FS\T$ is defined by $\FS\T(n) = \FS(n)\T$.
\item For commands, $\cbrk\T=\cbrk$ and $\cxynv\T = \cTxynv$.
\item For sequences of commands, $\emptyseq\T=\emptyseq$, and if $S'=\alpha\cc S$, then $S'\T=\alpha\T\cc S\T$.
\item Finally we extend $\T$ to sets of sequences: $\seqset{A}\T = \{S\T | S\in\seqset{A}\}$.
\end{itemize}
\end{mydef}

It is easy to see that for all objects where it is defined, $X\T\typeeq X$.
% $|\setf|$ is also $1$.
As representatives of filesystems, commands, sequences, etc.
are subsets of their non-representative counterparts, all
propositions proven so far apply to them as well.

@@

% TODO repeated
As whether a sequence of commands breaks a filesystem does not depend on the output values
of the commands, only the input and output types, we know that
\begin{mycor}\label{repr_works_is_same}
\begin{gather*}
\works{\seqset{A}} \Leftrightarrow \works{\seqset{A}\T} \\
\worksc{\seqset{A}}{\seqset{B}} \Leftrightarrow \worksc{\seqset{A}\T}{\seqset{B}\T}.
\end{gather*}
\end{mycor}

We are now ready to define the inverse of representative commands $\alpha\T$
which allows us to move parts of sequences between the
condition and consequence parts of $\workssign$.
This is based on the following observation.

% TODO partially bijective (?)
\begin{mylem}\label{repr_comm_inject}
Representative commands $\alpha\T$ are injections over representative filesystems ($\FS\T$)
except for the cases where they break a filesystem, that is,
\[ \forall\FS,\GS,\alpha: \FS\T\neq\GS\T \wedge \alpha\T(\FS\T)\neq\fsbroken \Longrightarrow \alpha\T(\FS\T) \neq \alpha\T(\GS\T). \]
\end{mylem}
\begin{proof}
This is trivial if $\GS\T=\fsbroken$, so we assume it is not so.
Let $\alpha\T = \cTxynv$.
If $\FS\T$ and $\GS\T$ have different values at any node apart from $n$, then this difference remains
even after applying $\cTxynv$ to $\FS\T$. The only remaining case therefore is where they have the same values
at all nodes except $n$.
As $\cTxynv(\FS\T)\neq\fsbroken$, we know $[\FS\T(n)]=[x]$.
We also know that $\GS\T(n)\neq\FS\T(n)$, and therefore, as there is a one-to-one correspondence between values and types,
$[\GS\T(n)]\neq[x]$.
This implies $\cTxynv(\GS\T)=\fsbroken$,
and again we have $\alpha\T(\FS\T) \neq \alpha\T(\GS\T)$.
\end{proof}

% TODO We introduced BREAK to deal with cases where commands
% are in effect not defined. Now we need to undo this 
% to see that repr commands are bijective.

% TODO Commands/sequences are in fact partial (one-to-one) functions, and/or partial bijections. See
% https://en.wikipedia.org/wiki/Inverse_semigroup#Origins
% https://en.wikipedia.org/wiki/Symmetric_inverse_semigroup
% https://en.wikipedia.org/wiki/Partial_function

\begin{mydef}[Inverse commands and sequences]
If based on \cref{repr_comm_inject}
we regard the representative command $\alpha\T$ a bijective function whose
domain is the subset of representative filesystems $\{\FS\T | \alpha\aFS\neq\fsbroken\}$,
then we can define the inverse of $\alpha\T$
for which we write $\alpha\T^{-1}$.
For any $\GS\T$ outside the image of the domain of $\alpha\T$ we define $\alpha\T^{-1}(\GS\T)$ to be $\fsbroken$.

We write ${S\T}^{-1}$ for the inverse of the sequence $S\T$, which consists of the inverses of the commands in $S\T$
in reverse order.
\end{mydef}

%% TODO
% Do we need <[x],[y],n>-1 = <[y],[x],n> ?
% Do we need to prove if a filesystem is not in the image, then a-1 will break it?
\begin{mylem}
For representative commands
\[ \cTxynv^{-1} = \cTyxnv. \]
\end{mylem}
TODO needed? -maybe for \cref{negneg_is_typeeq}

@@

For brevity, in the following lemma we use $B$, $B'$, $\seqset{A}$ and $\seqset{C}$ to denote
representative sequences and sets.

\begin{mylem}\label{r_invmove}
A common initial segment of a set of \emph{representative} sequences can be moved to the other side of $\worksmeqsign$ by inverting it:
\begin{gather*}
\worksm{\seqset{C}}{B\cc \seqset{A}}
\ \Longleftrightarrow \ \worksm{B^{-1}\cc \seqset{C}}{\seqset{A}} 
\ \wedge \  \worksm{\seqset{C}}{\{B\}}
\end{gather*}
\end{mylem}
\begin{proof}
% TODO Rewrite for equivalence, and swap A and C.
% This result is quite expected as we're mapping a subset statement using a bijection,
% but made more complicated a as B is not defined everywhere.

The proof for both propositions is illustrated by \cref{fig_invmove}.

@@

Unless $\works{\seqset{C}}$, both propositions are trivially true. Otherwise,
for the first proposition
we know $\works{B\cc\seqset{A}}$, which implies that $\Dom{\seqset{A}}$ intersects with $\Img{B}$ 
as otherwise $B\cc\seqset{A}$ would break all filesystems.
As the (representative) $B$ is a bijection between its domain and image,
we can use $B^{-1}$ to map this intersection back
onto the domain of $B$ and get $\Dom{B\cc\seqset{A}}$.
As $\worksc{B\cc \seqset{A}}{\seqset{C}}$, 
we know that $\Dom{\seqset{C}} \subseteq \Dom{B\cc\seqset{A}}$.
We now use $B$ to map both sides of this statement, which yields
$\Dom{B^{-1}\cc \seqset{C}} \subseteq \Dom{B^{-1}\cc B\cc\seqset{A}}$,
but the latter is a subseteq of $\Dom{\seqset{A}}$,
and so $\Dom{B^{-1}\cc \seqset{C}} \subseteq \Dom{\seqset{A}}$,
from which $\worksc{\seqset{A}}{B^{-1}\cc \seqset{C}}$.

For the second proposition, 
from $\worksc{B}{\seqset{C}}$ we know $\Dom{\seqset{C}} \subseteq \Dom{B}$.
Therefore we can project the whole of $\Dom{\seqset{C}}$ using the bijective representative sequence $B$
to get $\Dom{B^{-1}\cc\seqset{C}}$, and as $\worksc{\seqset{A}}{B^{-1}\cc \seqset{C}}$
we know
$\Dom{B^{-1}\cc\seqset{C}} \subseteq \Dom{\seqset{A}}$.
We use $B^{-1}$ to project both sides of this statement to get
$\Dom{B\cc B^{-1}\cc\seqset{C}} \subseteq \Dom{B\cc\seqset{A}}$.
But we also know that $\Dom{B\cc B^{-1}\cc\seqset{C}} = \Dom{\seqset{C}}$
as $\Dom{\seqset{C}} \subseteq \Dom{B}$,
and so $\Dom{\seqset{C}} \subseteq \Dom{B\cc\seqset{A}}$
from which $\worksc{B\cc \seqset{A}}{\seqset{C}}$.
\end{proof}

\begin{mylem}\label{indep_prefix_combine}
The combination of sequences with a common head and independent tails 
continues to work under the same conditions:
\[ \worksc{A\cc B}{\seqset{S}} \wedge \worksc{A\cc C}{\seqset{S}} \wedge B\indep C \Rightarrow \worksc{A\cc B\cc C}{\seqset{S}} \]
\end{mylem}
\begin{proof}
Based on \cref{repr_works_is_same} we have
$\worksc{A\T\cc B\T}{\seqset{S\T}}$ and $\worksc{A\T\cc C\T}{\seqset{S\T}}$,
and from \cref{r_invmove} we know
$\worksc{B\T}{{A\T}^{-1}\cc \seqset{S\T}}$ and $\worksc{C\T}{{A\T}^{-1}\cc \seqset{S\T}}$.
Using \cref{combine_independent_sequences} we get
$\worksc{B\T\cc C\T}{B\T,C\T}$,
and combining these using \cref{workschained}
we get
$\worksc{B\T\cc C\T}{{A\T}^{-1}\cc \seqset{S\T}}$.

From $\worksc{A\cc B}{\seqset{S}}$ we also know $\worksc{A}{\seqset{S}}$ and $\worksc{A\T}{\seqset{S\T}}$.
We can combine this and $\worksc{B\T\cc C\T}{{A\T}^{-1}\cc \seqset{S\T}}$ using
the second proposition of \cref{r_invmove} to get
$\worksc{A\T\cc B\T\cc C\T}{\seqset{S\T}}$, and from \cref{repr_works_is_same} we know
$\worksc{A\cc B\cc C}{\seqset{S}}$.
\end{proof}

>>>>

\begin{mydef}[Sets of sequences]
As we will frequently refer to sets of fequences,
we will use calligraphic letters (e.g. $\sqs{A}$, $\sqs{B}$, $\sqs{C}$ and $\sqs{S}$)
to denote such sets for brevity.
We will use $A\cc\sqs{B}$ to denote the set of sequences $\{A\cc B|B\in\sqs{B}\}$,
and $\Dom{\sqs{A}}$ to denote $\bigcap_{A\in\sqs{A}} \Dom{A}$.
\end{mydef}

% works operators
% ---------------

@@

In a sense, $A\worksmeqsign B$ is a weaker counterpart of $A\eqext B$, as the former
only requires that $B$ works where $A$ works, where the latter also requires
that their effect is the same.
It is easy to see that the following corollaries are true:

@@

\begin{mycor}\label{workschained}
% An inference rule in the algebra
$\worksmeqsign$ can be chained:
if $\sqs{B'}\subseteq\sqs{B}$, then
$\worksm{\sqs{A}}{\sqs{B}} \wedge \worksm{\sqs{B'}}{\sqs{C}} \Rightarrow \worksm{\sqs{A}}{\sqs{B}\cup\sqs{C}}$.
\end{mycor}

@@

% Type equivalence classes
% ------------------------

\noindent
For the proof of the correctness of the reconciliation algorithm,
we will also need a way of moving parts of sequences between the two
sides of the $\worksmeqsign$ relation.
We can do this based on the observation that commands are bijections
if we only distinguish between types in our objects.
To formalise this, we extend type equality ($\typeeq$)
to filesystems, commands and sequences of commands.
\begin{mydef}[Equivalence by type, extended: $\typeeq$]
$ $ % otherwise first list element starts on same line
\begin{itemize}
\item For two filesystems $\FS\typeeq\GS$ iff they share the same nodes $\setn$,
and for all $n\in\setn$, $\FS(n) \typeeq \GS(n)$.
\item For two commands $\cxynv\typeeq\czwmv$ iff 
$x\typeeq z$ and $y\typeeq w$
(that is, $[x]=[z]$ and $[y]=[w]$)
and $n=m$.
\item For two sequences of commands $\emptyseq\typeeq\emptyseq$, and
if $A'=\alpha\cc A$ and $B'=\beta\cc B$, then $A'\typeeq B'$ iff $\alpha\typeeq\beta$ and $A\typeeq B$.
\end{itemize}
\end{mydef}

@@

\begin{mydef}[$\T$]
We use $\T$ and write $a\T$ to denote the canonical mapping of 
values, filesystems, commands and sequences to the
representatives of their $\typeeq$-equivalence class.
\begin{itemize}
\item For values, $\vald\T=\vald$, $\valf\T=\fil$ and $\empt\T=\empt$.
\item For filesystems, $\FS\T$ is defined by $\FS\T(n) = \FS(n)\T$.
\item For commands, $\cbrk\T=\cbrk$ and $\cxynv\T = \cTxynv$.
\item For sequences of commands, $\emptyseq\T=\emptyseq$, and if $S'=\alpha\cc S$, then $S'\T=\alpha\T\cc S\T$.
\item Finally we extend $\T$ to sets of sequences: $\sqs{A}\T = \{S\T | S\in\sqs{A}\}$.
\end{itemize}
\end{mydef}

It is easy to see that for all objects where it is defined, $a\T\typeeq a$.
% $|\setf|$ is also $1$.
As representatives of filesystems, commands, sequences, etc.
are subsets of their non-representative counterparts, all
propositions proven so far apply to them as well.

@@

As the types of values determine whether a command breaks a filesystem,
we know $\Dom{A}=\Dom{A\T}$ and therefore $\Dom{\sqs{A}}=\Dom{\sqs{A}\T}$.
From this it follows that

\begin{mycor}\label{repr_works_is_same}
\begin{gather*}
\worksmnil{\sqs{A}} \Longleftrightarrow \worksmnil{\sqs{A}\T} \\
\worksm{\sqs{A}}{\sqs{B}} \Longleftrightarrow \worksm{\sqs{A}\T}{\sqs{B}\T}.
\end{gather*}
\end{mycor}

% Inverse commands
% ----------------

We are now ready to define the inverse of representative commands $\alpha\T$
which allows us to move parts of sequences between the
two sides of $\worksmeqsign$.
This is based on the following observation.

\begin{mylem}\label{repr_comm_inject}
Representative commands $\alpha\T$ are partial bijections over representative filesystems ($\FS\T$).
Because partial functions are trivially surjective when restricted to their images, this is
equivalent to saying that representative functions are injective except for cases
where they break a filesystem (they are not defined):
\[ \forall\FS,\GS,\alpha: \FS\T\neq\GS\T \wedge \alpha\T(\FS\T)\neq\fsbroken \Longrightarrow \alpha\T(\FS\T) \neq \alpha\T(\GS\T). \]
\end{mylem}
\begin{proof}
This is trivial if $\GS\T=\fsbroken$, so we assume it is not so.
Let $\alpha\T = \cTxynv$.
If $\FS\T$ and $\GS\T$ have different values at any node apart from $n$, then this difference remains
even after applying $\cTxynv$ to $\FS\T$. The only remaining case therefore is where they have the same values
at all nodes except $n$.
As $\cTxynv(\FS\T)\neq\fsbroken$, we know $[\FS\T(n)]=[x]$.
We also know that $\GS\T(n)\neq\FS\T(n)$, and therefore, as there is a one-to-one correspondence between values and types,
$[\GS\T(n)]\neq[x]$.
This implies $\cTxynv(\GS\T)=\fsbroken$,
and again we have $\alpha\T(\FS\T) \neq \alpha\T(\GS\T)$.
\end{proof}

% See
% https://en.wikipedia.org/wiki/Inverse_semigroup#Origins
% https://en.wikipedia.org/wiki/Symmetric_inverse_semigroup
% https://en.wikipedia.org/wiki/Partial_function

\begin{mydef}[Inverse commands and sequences]
We write $\alpha\T^{-1}$ for
the inverse of the representative command (partial bijection) $\alpha\T$.

We write ${S\T}^{-1}$ for the inverse of the sequence $S\T$, which consists of the inverses of the commands in $S\T$
in reverse order.
\end{mydef}
Per definition, for any filesystem $\FS\T$ outside the image of $\alpha\T$, the inverse $\alpha\T^{-1}$ is undefined.

\begin{mylem}
For representative commands
\[ \cTxynv^{-1} = \cTyxnv. \]
TODO needed? -maybe for \cref{negneg_is_typeeq}
\end{mylem}

@@

For brevity, in the following lemma we use $B$, $B'$, $\sqs{A}$ and $\sqs{C}$ to denote
representative sequences and sets.

\newcommand{\ia}{\sqs{A}}
\newcommand{\ic}{\sqs{C}}
\newcommand{\ibi}{B^{-1}}

\begin{mylem}\label{r_invmove}
A common initial segment of a set of \emph{representative} sequences can be moved 
to the other side of $\worksmeqsign$ by inverting it:
\begin{gather*}
\worksm{\ia}{B\cc\ic}
\ \Longleftrightarrow \ \worksm{\ibi\cc\ia}{\ic} 
\ \wedge \  \worksm{\ia}{\{B\}}
\end{gather*}
\end{mylem}
\begin{proof}
% This result is quite expected as we're mapping a subset statement using a bijection,
% but made more complicated a as B is not defined everywhere.

The proof is illustrated by \cref{fig_invmove}.

@@

Unless $\worksmnil{\ia}$, the propositions is trivially true. Otherwise,
for the left-to-right part of the proposition,
we know $\worksmnil{B\cc\ic}$, which implies that $\Dom{\ic}$ intersects with $\Img{B}$ 
as otherwise $B\cc\ic$ would break all filesystems.
As the (representative) $B$ is a bijection between its domain and image,
we can use $\ibi$ to map this intersection back
onto the domain of $B$ and get $\Dom{B\cc\ic}$.
As $\worksm{\ia}{B\cc\ic}$, 
we also know that $\Dom{\ia} \subseteq \Dom{B\cc\ic}$.
We now use $B$ to map both sides of this statement, which yields
$\Dom{B^{-1}\cc\ia} \subseteq \Dom{\ibi\cc B\cc\ic}$,
but the latter is a subset of $\Dom{\ic}$,
and so $\Dom{\ibi\cc\ia} \subseteq \Dom{\ic}$,
which is the same as $\worksm{\ibi\cc\ia}{\ic}$.

For the right-to-left part,
from $\worksm{\ia}{B}$ we know $\Dom{\ia} \subseteq \Dom{B}$.
Therefore we can project the whole of $\Dom{\ia}$ using the bijective representative sequence $B$
to get $\Dom{\ibi\cc\ia}$, and as $\worksm{\ibi\cc\ia}{\ic}$
we know
$\Dom{\ibi\cc\ia} \subseteq \Dom{\ic}$.
We use $\ibi$ to project both sides of this statement to get
$\Dom{B\cc\ibi\cc\ia} \subseteq \Dom{B\cc\ic}$.
But we also know that $\Dom{B\cc\ibi\cc\ia} = \Dom{\ia}$
as $\Dom{\ia} \subseteq \Dom{B}$,
and so $\Dom{\ia} \subseteq \Dom{B\cc\ic}$ which is the same as
$\worksm{\ia}{B\cc\ic}$.
\end{proof}

\begin{mylem}\label{indep_prefix_combine}
The combination of sequences with a common head and independent tails 
continues to work under the same conditions:
\[ \worksc{A\cc B}{\sqs{S}} \wedge \worksc{A\cc C}{\sqs{S}} \wedge B\indep C \Rightarrow \worksc{A\cc B\cc C}{\sqs{S}} \]
\end{mylem}
\begin{proof}
Based on \cref{repr_works_is_same} we have
$\worksc{A\T\cc B\T}{\sqs{S\T}}$ and $\worksc{A\T\cc C\T}{\sqs{S\T}}$,
and from \cref{r_invmove} we know
$\worksc{B\T}{{A\T}^{-1}\cc \sqs{S\T}}$ and $\worksc{C\T}{{A\T}^{-1}\cc \sqs{S\T}}$.
Using \cref{combine_independent_sequences} we get
$\worksc{B\T\cc C\T}{B\T,C\T}$,
and combining these using \cref{workschained}
we get
$\worksc{B\T\cc C\T}{{A\T}^{-1}\cc \sqs{S\T}}$.

From $\worksc{A\cc B}{\sqs{S}}$ we also know $\worksc{A}{\sqs{S}}$ and $\worksc{A\T}{\sqs{S\T}}$.
We can combine this and $\worksc{B\T\cc C\T}{{A\T}^{-1}\cc \sqs{S\T}}$ using
the second proposition of \cref{r_invmove} to get
$\worksc{A\T\cc B\T\cc C\T}{\sqs{S\T}}$, and from \cref{repr_works_is_same} we know
$\worksc{A\cc B\cc C}{\sqs{S}}$.
\end{proof}
